The Impact of Anthropogenic Forcing on ENSO Amplitude

### 2022-01-18 feedback/presentation_main.tex
stuff lol

% Created 2021-10-19 Tue 07:57
% Intended LaTeX compiler: pdflatex
\documentclass[little]{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\setcounter{secnumdepth}{1}
\author{Benjamin Goldman}
\date{\today}
\title{Feedback for main presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Feedback for main presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Feedback 1}
\label{sec:org1da6f23}
Date: \textit{<2021-09-26 Sun>}
Time: 20:20
Listener: Abby Goldman
\subsection*{Feedback}
\label{sec:org1d74604}
\begin{itemize}
\item lots of arm waving. SOOOOMUCH ARM WAVING!!! And very little breathing
\item make a list of “overhead” - make sure all concepts are well defined, in your notes, possibly in your slides
\item summary of what your many slides are — many different crunching of different data to circle around a process that we’re trying to under
\item insert heading slides, overall need a good deal more organization.
\item slide 1: break up into two or even 3?
\item Slide 2: Need to explain more generally what’s in the figure
\item Slide 3: climate simulation — is this method? If so, say so
\item Slide 4: ENSO and climate change slide — have “literature review” in heading.
\item slide about Mentor: do a better job explaining the interactive process between you and your mentor.
\item 3.4 index - whahah? I didn’t follow
\item Single forcing ensembles slide — not sure if all those arrows are working for you. Skip over single forcing setup, maybe?
\item use consistent language about the aerosols and their influence, considering that it’s a prediction
\end{itemize}

\section{Feedback 2}
\label{sec:org11974c6}
Date: \textit{<2021-09-30 Thu>}
Time: 8:20
Listener: Marcus Chung, Sean O'Donnell, Jimena Perez
\subsection*{Feedback:}
\label{sec:orgdbbe7e5}
\begin{itemize}
\item Figure text is too small
\item Caption text can be smaller to make room
\item Say some transition for the conclusion heading slide
\item Clarify difference between El Niño and La Niña
\begin{itemize}
\item overall make intro 2 more organized
\end{itemize}
\item Good understanding and well communicated
\item Are some slides too empty-looking?
\item Say fewer words because it is over time and talking too fast
\item Use note cards
\item Periods should be consistent.
\end{itemize}

\section{Feedback 3}
\label{sec:org52baac0}
Date: \textit{<2021-09-30 Thu>}
Time: 16:00
Listener: Michael Goldman
\subsection*{Feedback:}
\label{sec:orgf282c69}
\begin{itemize}
\item Make a few bullets clearer
\item Good tone
\item Slow down pacing
\item Introduction is slightly unclear
\item Use note cards
\end{itemize}

\section{Feedback 4}
\label{sec:org585437d}
Date: \textit{<2021-10-17 Sun>}
Time: 17:12
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgeffe568}
\begin{itemize}
\item Make figures bigger and make captions smaller
\item Love how detailed the content is
\item Great application slide
\item Make sure you explain your results in the context of your study, think big picture
\item Intro slide 2 is great we’ll explained
\end{itemize}
\end{document}

>>>>

% Created 2022-01-06 Thu 19:00
% Intended LaTeX compiler: pdflatex
\documentclass[little]{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\setcounter{secnumdepth}{1}
\author{Benjamin Goldman}
\date{\today}
\title{Feedback for main presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Feedback for main presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Feedback 1}
\label{sec:orgb0ce4b1}
Date: \textit{<2021-09-26 Sun>}
Time: 20:20
Listener: Abby Goldman
\subsection*{Feedback}
\label{sec:orgeae9fd0}
\begin{itemize}
\item lots of arm waving. SOOOOMUCH ARM WAVING!!! And very little breathing
\item make a list of “overhead” - make sure all concepts are well defined, in your notes, possibly in your slides
\item summary of what your many slides are — many different crunching of different data to circle around a process that we’re trying to under
\item insert heading slides, overall need a good deal more organization.
\item slide 1: break up into two or even 3?
\item Slide 2: Need to explain more generally what’s in the figure
\item Slide 3: climate simulation — is this method? If so, say so
\item Slide 4: ENSO and climate change slide — have “literature review” in heading.
\item slide about Mentor: do a better job explaining the interactive process between you and your mentor.
\item 3.4 index - whahah? I didn’t follow
\item Single forcing ensembles slide — not sure if all those arrows are working for you. Skip over single forcing setup, maybe?
\item use consistent language about the aerosols and their influence, considering that it’s a prediction
\end{itemize}

\section{Feedback 2}
\label{sec:org3a9f9f0}
Date: \textit{<2021-09-30 Thu>}
Time: 8:20
Listener: Marcus Chung, Sean O'Donnell, Jimena Perez
\subsection*{Feedback:}
\label{sec:orgd187f3b}
\begin{itemize}
\item Figure text is too small
\item Caption text can be smaller to make room
\item Say some transition for the conclusion heading slide
\item Clarify difference between El Niño and La Niña
\begin{itemize}
\item overall make intro 2 more organized
\end{itemize}
\item Good understanding and well communicated
\item Are some slides too empty-looking?
\item Say fewer words because it is over time and talking too fast
\item Use note cards
\item Periods should be consistent.
\end{itemize}

\section{Feedback 3}
\label{sec:org304f650}
Date: \textit{<2021-09-30 Thu>}
Time: 16:00
Listener: Michael Goldman
\subsection*{Feedback:}
\label{sec:org2526349}
\begin{itemize}
\item Make a few bullets clearer
\item Good tone
\item Slow down pacing
\item Introduction is slightly unclear
\item Use note cards
\end{itemize}

\section{Feedback 4}
\label{sec:org661561a}
Date: \textit{<2021-10-17 Sun>}
Time: 17:12
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgfb0e433}
\begin{itemize}
\item Make figures bigger and make captions smaller
\item Love how detailed the content is
\item Great application slide
\item Make sure you explain your results in the context of your study, think big picture
\item Intro slide 2 is great we’ll explained
\end{itemize}

\section{Feedback 5}
\label{sec:org4ee2cd0}

Date: \textit{<2022-01-06 Thu>}
Time: 09:20
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgcae29f9}
\begin{itemize}
\item Put Student vs mentor role earlier in presentation
\item Slow down speech
\item Write a script for wavelet slide
\item Don't admit you don't understand it
\item No monospace for urls
\item Clean up global warming slide
\end{itemize}
\end{document}

### 2022-01-18 presentations/presentation/presentation.tex
stuff lol

\begin{frame}{El Niño (ENSO)}
  \begin{figure}
    \begin{columns}
      \column{0.5\textwidth}
      \begin{itemize}
      \item Warming and cooling of the Pacific Ocean.
      \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
      \item May be affected by climate change.
      \end{itemize}
      \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
      \column{0.5\textwidth}
      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
    \end{columns}
  \end{figure}
\end{frame}

@@

\begin{frame}{Gap and Questions}
  \begin{itemize}
  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
  \end{itemize}
  \begin{description}
  \item[What?] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
  \item[Why?] Is the predicted increase (or decrease) due to human activities?
  \item[How?] What processes are causing greenhouse gasses and aerosols to affect ENSO?
  \end{description}
\end{frame}

\begin{frame}{Model Setup (Data)}
  \begin{itemize}
  \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
  \item Predicts climate over 21st century with global warming.
  \item 40-50 simulations per ensemble.
  \item Control simulation with pre-1850 forcing levels.
  \item Single forcing ensembles that represent influence of single factor.
  \end{itemize}
\end{frame}

\begin{frame}{Measuring ENSO Intensity}
  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
    \node [data] (input) {
      \begin{varwidth}{1.5in}
        External forcing
      \end{varwidth}};
    \node [process] (model) [right of=input] {
      \begin{varwidth}{\linewidth}
        CESM1\\and\\CESM2
      \end{varwidth}};
    \node [data] (output) [right of=model] {
      \begin{varwidth}{1.2in}
        Sea surface temperature (SST)
      \end{varwidth}};
    \node [data] (nino) [below of=input] {
      \begin{varwidth}{1.4in}
        Niño 3.4 index
      \end{varwidth}};
    \node [process] (variance) [right of=nino] {
      \begin{varwidth}{1.4in}
        Windowed\\variance
      \end{varwidth}};
    \node [data] (intensity) [right of=variance] {
      \begin{varwidth}{1.4in}
        ENSO intensity
      \end{varwidth}};
    \draw [->] (input) to (model);
    \draw [->] (model) to (output);
    \draw [->] (output) to [out = 270, in = 90] (nino);
    \draw [->] (nino) to (variance);
    \draw [->] (variance) to (intensity);
  \end{tikzpicture}
\end{frame}

\begin{frame}{ENSO is Becoming Stronger}
  \begin{columns}
    \column{0.4\textwidth}
    \begin{itemize}
    \item Increase in ENSO intensity in both ensembles.
    \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
    \end{itemize}
    \column{0.6\textwidth}
    \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{ff_compare}
  \end{columns}
\end{frame}

@@

\begin{frame}{Wavelet Analysis}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item Separate ENSO record into changes in period over time.
    \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
    \item In CESM1, increase in ENSO intensity is mainly strengthening of longer-period cycle.
    \item In CESM2, longer-period ENSO weakens after 2025.
    \end{itemize}
    \column{0.5\textwidth}
    \myfig{wavelet3.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
  \end{columns}
\end{frame}

@@

\begin{frame}{Limitations and Applications}
  Limitations:
  \begin{itemize}
  \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
  \item CESM may contain biases.
  \end{itemize}
  Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
\end{frame}

>>>>

\begin{frame}{El Niño (ENSO)}
  \begin{figure}
    \begin{columns}
      \column{0.5\textwidth}
      \begin{itemize}
      \item ENSO = El Niño/Southern Oscillation
      \item Warming and cooling of the Pacific Ocean.
      \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
      \item May be affected by climate change.
      \end{itemize}
      \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
      \column{0.5\textwidth}
      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
    \end{columns}
  \end{figure}
\end{frame}

@@

\begin{frame}{Gap and Questions}
  \alert{Gap}
  \begin{itemize}
  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
  \end{itemize}
  \alert{Questions}
  \begin{description}
  \item[\alert{What?}] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
  \item[\alert{Why?}] Is the predicted increase (or decrease) due to human activities?
  \item[\alert{How?}] What processes are causing greenhouse gasses and aerosols to affect ENSO?
  \end{description}
\end{frame}

\section{Methods and Results}

\begin{frame}{Model Setup (Data)}
  \begin{itemize}
  \item Community Earth System Model (CESM) Versions 1 and 2 \citep{kay2015community} \citep{danabasoglu2020community}.
  \item Predicts climate over 21st century with global warming.
  \item 40-50 simulations per ensemble.
  \item Control simulation with pre-1850 forcing levels.
  \item Single forcing ensembles that represent influence of single factor.
  \end{itemize}
\end{frame}

\begin{frame}{Measuring ENSO Intensity}
  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
    \node [process] (input) {
      \begin{varwidth}{1.2in}
        External forcing records and predictions
      \end{varwidth}};
    \node [process] (model) [right of=input] {
      \begin{varwidth}{1.2in}
        CESM1 and CESM2 climate models
      \end{varwidth}};
    \node [process] (output) [right of=model] {
      \begin{varwidth}{1.2in}
        Sea surface temperature (SST)
      \end{varwidth}};
    \node [process] (nino) [below of=input] {
      \begin{varwidth}{1.2in}
        Mean tropical Pacific temperature
      \end{varwidth}};
    \node [process] (variance) [right of=nino] {
      \begin{varwidth}{1.0in}
        Windowed variance: Mean variance over every 20 years.
      \end{varwidth}};
    \node [process] (intensity) [right of=variance] {
      \begin{varwidth}{1.2in}
        Mean and standard error
      \end{varwidth}};
    \draw [->] (input) to (model);
    \draw [->] (model) to (output);
    \draw [->] (output) to [out = 270, in = 90] (nino);
    \draw [->] (nino) to (variance);
    \draw [->] (variance) to (intensity);
  \end{tikzpicture}
\end{frame}

\begin{frame}{ENSO is Becoming Stronger}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item Increase in ENSO intensity in both ensembles. (Exceeds 2 standard errors)
    \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
    \end{itemize}
    \column{0.5\textwidth}
    \begin{figure}
      \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\textwidth]{figures/ff_compare.pdf}};
        \coordinate (a) at (0.5, 1.5);
        \node [red] (b) at (1.0, 1.0) {
          \begin{varwidth}{0.6in}
            Increasing variance
          \end{varwidth}};
        \coordinate (c) at (3.0, 1.5) {};
        \draw[red, |-|] (a) to (c);

        \coordinate (d) at (3.2, 1.5);
        \node [red] (e) at (3.5, 1.0) {
          \begin{varwidth}{0.6in}
            Decreasing variance
          \end{varwidth}};
        \coordinate (f) at (4.3, 1.5) {};
        \draw[red, |-|] (d) to (f);
      \end{tikzpicture}
      \caption{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}
      \label{ff_compare}
    \end{figure}
  \end{columns}
\end{frame}

@@

\begin{frame}{Wavelet Analysis}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item Separate ENSO record into changes in period over time.
    \item In CESM1, increase in ENSO intensity is mainly strengthening of longer-period cycle.
    \item In CESM2, longer-period ENSO weakens after 2025.
    \item Indicates that longer frequency bands are more susceptible to climate change.
    \end{itemize}
    \column{0.5\textwidth}
    \myfig{wavelet3.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
  \end{columns}
\end{frame}

@@

\begin{frame}{Limitations and Applications}
  Limitations:
  \begin{itemize}
  \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
  \item CESM may contain biases.
  \item Models are only an approximation of the Earth's actual climate.
  \end{itemize}
  Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
\end{frame}
